\section{Binary Installation}
There are several required elements to use the binary installation.

\begin{itemize}
\item LLVM - obtain version 1.5 from www.llvm.org
\item LLVM cfrontend - obtain version 1.5 from www.llvm.org
\item antlr.jar - obtain 2.7.3 or later from www.antlr.org
\item java-getopt.jar - obtain version 1.11 or later from \\
http://www.urbanophile.com/arenn/hacking/download.html
\item Java - obtain from java.sun.com (tested with 1.4.2)
\item gcc, unix environment, etc.
\item Graphviz (dotty and friends) from www.graphviz.org can be useful, but is not required.
\end{itemize}

First obtain LLVM and compile LLVM.  The following works on RHEL 3.0u5
with tcsh, and detailed instructions are available on www.llvm.org.
If you prefer bash, use the proper commands for setting variables
(e.g., export BLAH=SOME\_VALUE and export
PATH=\$PATH:NEW\_PATH\_VALUE, etc.).

\begin{verbatim}
> tar xzf llvm-1.5.tar.gz
> tar xzf cfrontend-1.5.i686-redhat-linux-gnu.tar.gz

#
# must have java
#
> which java
>
> setenv TOP `pwd`

> cd cfrontend/x86/
> ./fixheaders
> cd ../..

> set path = ( $path $TOP/cfrontend/x86/llvm-gcc/bin/ )

> mkdir llvm-obj
> cd llvm-obj

> ../llvm/configure
> make
# wait 

> cd ..
> set path = ( $path $TOP/llvm-obj/Debug/bin/ )

\end{verbatim}

Now LLVM should be built.  This will build the Debug version of LLVM and anything
else may not work without changing some of the scripts provided.

Next, unpack the Trident files and execute the following:\footnote{This is a continuation of the above
and uses the same TOP variable.  It could be modified to support a different location than parallel
with LLVM.}

\begin{verbatim}
> tar xzf dist.tgz

# build llv
# 
> cd trident/llv-src
> ./mkconf.sh --src=$TOP/llvm
> ./configure --with-llvmsrc=$TOP/llvm/ --with-llvmobj=$TOP/llvm-obj/

> make
> cp Debug/bin/llv ../bin
> strip ../bin/llv
> chmod a+x ../bin/llv

> cd ..
> set path = ($path $TOP/trident/bin )

> setenv TRIDENT_TOP $TOP/trident

# copy jars to $TRIDENT_TOP/lib
#
> cp somewhere/antlr.jar lib/
> cp somewhere/java-getopt.jar lib/
\end{verbatim}

Now, you should have all the tools necessary to run Trident.  Test the 
compiler using the following:

\begin{verbatim}
cd examples/conditional
tcc -t vhdl conditional_a.c run

less conditional_a.vhd
\end{verbatim}

This will generate a VHDL file for the conditional\_a.c example.


